---
title: "Week 9 Redo"
format: html
editor: visual
---

## Setup

```{r}
library(tidyverse)
library(gssr)
library(boot)
theme_set(theme_light(base_family = "Optima"))

gss18 <- gss_get_yr(2018) 

d <- gss18 |> 
  select(attend, polviews, cappun) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  mutate(
    weekly = if_else(attend >= 7, 1L, 0L),
    conservative = if_else(polviews >= 5, 1L, 0L),
    cappun = if_else(cappun == 1, 1L, 0L)
    ) |> 
  drop_na() |> 
  select(cappun, weekly, conservative) 
```

## 9.1.1 Exercise

**The following contingency table shows the association between 2 variables: cappun and weekly.**

**Express the relationship between cappun (Y) and weekly (X) as a linear probability model.**

```{r}
crosstab1 <- table(cp = d$cappun, weekly = d$weekly)
crosstab1
 
# to view row percentages, set margin = 1 (good if interested in weekly as y, cappun as x)
prop.table(crosstab1, margin = 1)

# to view column percentages, set margin = 2 (good if interested in cappun as y, weekly as x)
prop.table(crosstab1, margin = 2)

# to get alpha, what is the probability of cappun = 1 if weekly = 0
alpha <- prop.table(crosstab1, margin = 2)[2,1] # second row, first column

# to get beta, subtract alpha from the probability of cappun = 1 if weekly = 1
beta <- prop.table(crosstab1, margin = 2)[2,2] - alpha

# confirm answer using lm()
lm(d$cappun ~ d$weekly)
```

**Express the relationship between cappun (Y) and weekly (X) as a logistic regression model**

```{r}
prop.table(crosstab1, margin = 2)

# logistic regression, so we want log-odds

alpha_logit <- log(prop.table(crosstab1, margin = 2)[2,1]/(1-prop.table(crosstab1, margin = 2)[2,1]))

beta_logit <- log(prop.table(crosstab1, margin = 2)[2,2]/(1-prop.table(crosstab1, margin = 2)[2,2]))-alpha_logit

# confirm answer using glm()
glm(d$cappun ~ d$weekly, family = "binomial")
```

## 9.1.2 Exercise

**The following table adds a third variable to the previous contingency table: conservative.**

```{r}
crosstab2 <- table(cp = d$cappun, weekly = d$weekly, conservative = d$conservative)
crosstab2

## column percentages
prop.table(crosstab2, margin = c(2, 3))

```

**Express the relationship between `cappun` (** Y **), `weekly` (** X1 **), and `conservative` (** X2 **) as a linear probability model.**

```{r}
# changing 2x2x2 table to be wider
as_tibble(crosstab2) |> 
  mutate(cp = if_else(cp == "1", "favors", "opposes")) |> 
  pivot_wider(names_from = cp, values_from = n)

# to get alpha, what is the probability of cappun = 1 if weekly = 0 and conservative = 0?
alpha <- prop.table(crosstab2, margin = c(2, 3))[2,1,1]

# to get beta1, subtract alpha from the probability of cappun = 1 if weekly = 1 and conservative = 0
beta_1<- prop.table(crosstab2, margin = c(2, 3))[2,2,1] - alpha

# to get beta2, subtract alpha from the probability of cappun = 1 if weekly = 0 and conservative = 1
beta_2<- prop.table(crosstab2, margin = c(2, 3))[2,1,2] - alpha

# to get beta3, subtract alpha, beta1, and beta2, from the probability of cappun = 1 if weekly = 1 and conservative = 1
beta_3 <- prop.table(crosstab2, margin = c(2, 3))[2,2,2] - beta_1- beta_2-alpha

# checking answers
lm(d$cappun ~ d$weekly + d$conservative + d$weekly*d$conservative)
```

**Express the relationship between `cappun` (** Y **), `weekly` (** X1 **), and `conservative` (** X2 **) as a logistic regression model.**

```{r}
alpha_logit <- log(prop.table(crosstab2, margin = c(2, 3))[2,1,1]/(1-prop.table(crosstab2, margin = c(2, 3))[2,1,1]))

beta1_logit <- log(prop.table(crosstab2, margin = c(2, 3))[2,2,1]/(1-prop.table(crosstab2, margin = c(2, 3))[2,2,1]))-alpha_logit

beta2_logit <- log(prop.table(crosstab2, margin = c(2, 3))[2,1,2]/(1-prop.table(crosstab2, margin = c(2, 3))[2,1,2]))-alpha_logit

beta3_logit <- log(prop.table(crosstab2, margin = c(2, 3))[2,2,2]/(1-prop.table(crosstab2, margin = c(2, 3))[2,2,2])) - beta2_logit - beta1_logit - alpha_logit

# checking answers
glm(d$cappun ~ d$weekly + d$conservative + d$weekly*d$conservative, family = "binomial")
```
