---
title: "Exercise_07"
author: "Lev Kolinski"
format: html
editor: visual
---

## 7.1.1

```{r include = FALSE}
library(tidyverse, warn.conflicts = FALSE)
library(gssr)
library(janitor)
library(infer)
gss18 <- gss_get_yr(2018) 

d <- gss18 |> 
  select(attend, polviews) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  drop_na() |> 
  mutate(weekly = if_else(attend >= 7, 1L, 0L),
         attends = if_else(weekly == 1, "Weekly", "Not Weekly"),
         conservative = if_else(polviews >= 5, 1L, 0L),
         polview_chr = if_else(conservative==1, "Conservative", "Not Conservative")) |> 
  relocate(attends,weekly,polview_chr,conservative)
```

1.  Probability that someone attends religious services weekly

```{r}
tabyl(d, weekly)
total_sample<-nrow(d)

prob_attends_church<-497/total_sample*100
prob_attends_church
```

Pr(W) = **22.24%**

2.  Probability that someone does NOT attend religious services weekly:

```{r}
100-prob_attends_church
```

Pr(W^c^) = **77.76%**

3.  Probability that a person is conservative given that they attend religious services weekly:

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "row") |> 
  adorn_pct_formatting(digits = 2)
```

Pr(C \| W) = **48.29%**

4.  Probability that a person attends services weekly given that they are conservative

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "col") |>
  adorn_pct_formatting(digits = 2)
```

Pr(W \| C) = **32.65%**

5.  Probability that a person is conservative given that they do NOT attend services weekly

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "row") |>
  adorn_pct_formatting(digits = 2)
```

Pr(C \| W^c^) = **28.48%**

6.  Probability that a person attends services weekly AND is conservative

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "all") |>
  adorn_pct_formatting(digits = 2)
```

Pr(W and C) = **10.74%**

7.  Probability that a person does NOT attend services weekly AND is NOT conservative

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "all") |>
  adorn_pct_formatting(digits = 2)
```

Pr(W^c^ and C^c^ ) = **55.62%**

8.  If independent, what would be the *joint* probability of W and C?

```{r}
chisq<-tabyl(d, attends, polview_chr) |> 
  chisq.test()
chisq$expected |> 
  adorn_percentages(denominator = "all")
```

The joint probability of W and C would be **7.31%**

## 7.1.2

Selecting new variables: marital status (M) and whether or not respondent has kids (K)

```{r}
d2 <- gss18 |> 
  select(marital, childs) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  drop_na() |> 
  mutate(married = if_else(marital == 1, 1L, 0L),
         mar_status_chr = if_else(marital == 1, "Married", "Not Married"),
         children = if_else(childs >= 1, 1L, 0L),
         has_children_chr = if_else(children==1, "Has kids", "No kids")) |>
  relocate(married,mar_status_chr,children,has_children_chr)

```

Contingency table:

```{r}
tabyl(d2, mar_status_chr, has_children_chr)
```

1.  Probability that someone is married: Pr(M)

```{r}
tabyl(d2, mar_status_chr)

prob_married<-996/(996+1347)*100
prob_married
```

Pr(M) = **42.51%**

2.  Probability that someone is NOT married: Pr(M^c^)

```{r}
100-prob_married
```

Pr(M^c^) = **57.49%**

3.  Probability that someone has children given that they are married: Pr(K \| M)

```{r}
tabyl(d2, mar_status_chr, has_children_chr) |> 
  adorn_percentages(denominator = "row") |>
  adorn_pct_formatting(digits = 2)
```

Pr(K \| M) = **85.44%**

4.  Probability that someone is married given that they have kids: Pr(M \| K)

```{r}
tabyl(d2, mar_status_chr, has_children_chr) |> 
  adorn_percentages(denominator = "col") |>
  adorn_pct_formatting(digits = 2)
```

Pr(M \| K) = **50.65%**

5.  Probability that someone has kids given that they are NOT married: Pr(K \| M^c^)

```{r}
tabyl(d2, mar_status_chr, has_children_chr) |> 
  adorn_percentages(denominator = "row") |>
  adorn_pct_formatting(digits = 2)
```

Pr(K \| M^c^) = **61.54%**

6.  Probability that someone is married and has kids: Pr(M and K)

```{r}
tabyl(d2, mar_status_chr, has_children_chr) |> 
  adorn_percentages(denominator = "all") |>
  adorn_pct_formatting(digits = 2)
```

Pr(M and K) = **36.32%**

7.  Probability that someone is NOT married and does NOT have kids: Pr(M^c^ and K^c^)

```{r}
tabyl(d2, mar_status_chr, has_children_chr) |> 
  adorn_percentages(denominator = "row") |>
  adorn_pct_formatting(digits = 2)
```

Pr(M^c^ and K^c^) = **38.46%**

8.  If M and K are independent, what would be the joint probability of M and K?

```{r}
chisq<-tabyl(d2, mar_status_chr, has_children_chr) |> 
  chisq.test()
chisq$expected |> 
  adorn_percentages(denominator = "all")
```

The joint probability of M and K would be **30.48%**

## 7.2.1

Compare Pr(W\|C) and Pr(W\|C^c^) using the following summary statistics:

1.  Difference in Probabilities

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "col") |>
  adorn_pct_formatting(digits = 2)
prob_weekly_given_conservative<-.3265
prob_weekly_given_not_conservative<-.1713

prob_weekly_given_conservative-prob_weekly_given_not_conservative
```

There is a **15.52% point difference in probabilities** between the two groups: probability of going to services weekly among those who are conservative AND probability of going to services weekly among those who are not conservative.

2.  Relative risk ratio

```{r}
prob_weekly_given_conservative/prob_weekly_given_not_conservative
```

People who are conservative have a probability of attending religious services weekly about **1.9X greater** than those who are not conservative.

3.  Odds ratio

```{r}
tabyl(d, attends, polview_chr)
# odds that person who is conservative attends weekly services
odds_weekly_conservative<-240/495

# odds that a person who is NOT conservative attends weekly services
odds_weekly_not_conservative<-257/1243

odds_ratio<-odds_weekly_conservative/odds_weekly_not_conservative
odds_ratio
```

Conservatives' odds of attending weekly religious services are **2.3X greater** than those who are not conservative

4.  Log odds ratio

```{r}
log(odds_ratio)
```

When compared to non-conservative, the log-odds ratio of conservatives attending weekly religious services is **about 0.85**

## 7.2.2

1.  Draw a sampling distribution for the relative risk ratio between Pr(W\|C) and Pr(W\|C^c^).

```{r}
d <- gss18 |> 
  select(attend, polviews) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  drop_na()

d <- d |> 
  mutate(weekly = if_else(attend >= 7, 1L, 0L),
         conservative = if_else(polviews >= 5, 1L, 0L)) |> 
  drop_na() 

boot_rr <- d |> 
  mutate(conservative_cat = if_else(conservative == 1L, "conservative", "other")) |> 
  mutate(weekly_cat = if_else(weekly == 1L, "weekly", "not_weekly")) |> 
  specify(weekly_cat ~ conservative_cat, success="weekly") |> 
  generate(reps = 1e4, type = "bootstrap") |> 
  calculate(stat = "ratio of props", order = c("conservative", "other"))

ci_rr <- get_confidence_interval(boot_rr)

boot_rr |> 
  visualize() +
  shade_ci(ci_rr)+
  ggtitle("relative risk ratio: Pr(W|C) and Pr(W|C^c)")
```

2.  Draw a sampling distribution for the odds ratio between Pr(W\|C) and Pr(W\|C^c^).

```{r}
boot_or <- d |> 
  mutate(conservative_cat = if_else(conservative == 1L, "conservative", "other")) |> 
  mutate(weekly_cat = if_else(weekly == 1L, "weekly", "not_weekly")) |> 
  specify(weekly_cat ~ conservative_cat, success="weekly") |> 
  generate(reps = 1e4, type = "bootstrap") |> 
  calculate(stat = "odds ratio", order = c("conservative", "other"))

ci_or <- get_confidence_interval(boot_or)

boot_or |> 
  visualize() +
  shade_ci(ci_or)+
  ggtitle("odds ratio: Pr(W|C) and Pr(W|C^c)")
```

## 7.2.3

Compare Pr(M\|K) and Pr(M\|K^c^) using the following summary statistics:

1.  Difference in Probabilities

```{r}
tabyl(d2, mar_status_chr, has_children_chr) |>
  adorn_percentages(denominator = "col") |>
  adorn_pct_formatting(digits = 2)
prob_married_given_kids<-.5065
prob_married_given_no_kids<-.2187

prob_married_given_kids-prob_married_given_no_kids
```

There is a **28.78% point difference in probabilities** between the two groups: probability of being married among those who have kids and probability of being married among those who do NOT have kids.

2.  Relative risk ratio

```{r}
prob_married_given_kids/prob_married_given_no_kids
```

People who have kids have a probability of being married about **2.32X greater** than those who do not have kids.

3.  Odds ratio

```{r}
tabyl(d2, mar_status_chr, has_children_chr)
# odds that person who has kids is married
odds_married_yes_kids<-851/829

# odds that a person who does NOT have kids is married
odds_married_no_kids<-145/518

odds_ratio_married_kids<-odds_married_yes_kids/odds_married_no_kids
odds_ratio_married_kids
```

Respondents with kids' odds of being married is about **3.67X greater** than those who do not have kids.

4.  Log odds ratio

```{r}
log(odds_ratio_married_kids)
```

When compared to respondents without kids, the log-odds ratio of respondents with kids being married is **about 1.30.**

## 7.2.4

1.  Draw a sampling distribution for the relative risk ratio between Pr(M\|K) and Pr(M\|K^c^).

```{r}
boot_rr_d2 <- d2 |> 
  specify(mar_status_chr ~ has_children_chr, success="Married") |> 
  generate(reps = 1e4, type = "bootstrap") |> 
  calculate(stat = "ratio of props", order = c("Has kids", "No kids"))

ci_rr_d2 <- get_confidence_interval(boot_rr_d2)

boot_rr_d2 |> 
  visualize() +
  shade_ci(ci_rr_d2)+
  ggtitle("relative risk ratio: Pr(M|K) and Pr(M|K^c)")
```

2.  Draw a sampling distribution for the odds ratio between Pr(M\|K) and Pr(M\|K^c^).

```{r}
boot_or_d2 <- d2 |> 
  specify(mar_status_chr ~ has_children_chr, success="Married") |> 
  generate(reps = 1e4, type = "bootstrap") |> 
  calculate(stat = "odds ratio", order = c("Has kids", "No kids"))

ci_or_d2 <- get_confidence_interval(boot_or_d2)

boot_or_d2 |> 
  visualize() +
  shade_ci(ci_or_d2)+
  ggtitle("odds ratio: Pr(M|K) and Pr(M|K^c)")
```
